---
title: "NC State Senate Facebook Posts (Part 2)"
author: "Ryan Wesslen"
date: "Jan 4, 2017"
output: html_document
---
  
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

### STM Exploration

#### Topic Summary

First, let's load our STM results from Part 1.

```{r}
load("~/Dropbox (UNC Charlotte)/NCStateSenateFacebook/data/stmFit.RData")
load("~/Dropbox (UNC Charlotte)/NCStateSenateFacebook/data/out.RData")
```

Let's explore the size of the topics by their topic proportions.

```{r fig.width = 8, fig.height=8}
library(stm)
plot.STM(stmFit, type = "summary", xlim = c(0,.14), n = 5, #labeltype = "frex",
         main = "NC State Senators' Topics on Facebook", text.cex = 0.8)
```

I've assigned labels to the topics based on my interpretation of the word-topic probabilities (see next section).

```{r results="hide"}
topicNames <- labelTopics(stmFit)
k <- 40
topic <- data.frame(
  topicnames = c("Press Conference",
                 "Local Government",
                 "Religious Freedom",
                 "#WeAreNotThis",
                 "Positive Outlook",
                 "Pat McCrory",
                 "Legislation",
                 "Gerrymandering",
                 "Supreme Court",
                 "Voter ID/Fraud",

                 "Presidential Election",
                 "Campaign Support",
                 "Health Care",
                 "Congressional Elections",
                 "North Carolina",
                 "HB2",
                 "Constituent Newsletter",
                 "Sen Van Duyn Posts",
                 "God, Family, Freedom",
                 "Teacher Pay",

                 "#NCPOL and #NCGA",
                 "Hurricane Matthew",
                 "Redistricting",
                 "Taxes",
                 "Congratulations",
                 "Bathroom Safety",
                 "Economy/Jobs",
                 "Student/Women's Issues",
                 "Hillary Clinton",
                 "Church",

                 "Civil Rights",
                 "Energy Tax Credits",
                 "Roy Cooper",
                 "Get Out the Vote",
                 "Gun Violence",
                 "Berger Press Releases",
                 "Public Assistance",
                 "Town Hall Events",
                 "Holiday Wishes",
                 "Conservative Values"),
  TopicNumber = 1:k,
  TopicProportions = colMeans(stmFit$theta))
```

#### Interpretating the Topics

```{r fig.width=8}
par(mfrow = c(4,2),mar = c(1, 1, 2, 1))
for (i in 1:k){
  plot.STM(stmFit, type = "labels", n = 15, 
           topics = i, main = paste0(topic$topicnames[i]," - Raw Probabilities"),
           width = 55)
  plot.STM(stmFit, type = "labels", n = 15, 
           topics = i, main = paste0(topic$topicnames[i]," - FREX"), 
           labeltype = "frex", width = 55)
}
```

```{r}
#shortdoc <- substr(subset(fbData$ID,out$meta$ID),1,200))
#thoughts3 <- findThoughts(stmFit, texts = shortdoc, n = 2, topics = 2)
```

```{r fig.height=6}
par(mfrow = c(1,1),mar = c(2, 2, 2, 2))
topicQuality(stmFit,
             documents = out$documents, 
             main = "Topic Interpretability: Exclusivity and Semantic Coherence")
```

```{r fig.height=10, fig.width=8}
prep <- estimateEffect(1:k ~ Party + s(Time), stmFit, meta = out$meta, uncertainty = "Global")

Result <- plot.estimateEffect(prep, "Party", method = "difference", 
                              cov.value1 = "Democratic", cov.value2 = "Republican", 
                              verbose.labels = F, 
                              ylab = "Expected Difference in Topic Probability by Party \n (with 95% Confidence Intervals)", 
                              xlab = "More Likely Republican                         Not Significant                          More Likely Democratic",
                              main = "Effect of Party on Topic Prevelance for Facebook Posts of NC State Senators",
                              xlim = c(-0.08,0.08))

# order based on Expected Topic Proportion
rank = order(unlist(Result$means))
topic <- topic[rank,]

par(mfrow = c(1,1),mar = c(6, 6, 4, 4))
STMresults <- plot.estimateEffect(prep, "Party", method = "difference", cov.value1 = "Democratic", 
                                  cov.value2 = "Republican", 
                                  topics = topic$TopicNumber,
                                  verbose.labels = F, 
                                  ylab = "Expected Difference in Topic Probability by Party \n (with 95% Confidence Intervals)", 
                                  labeltype = "custom",
                                  xlab = "More Likely Republican                         Not Significant                          More Likely Democratic",
                                  custom.labels  = topic$topicnames, 
                                  main = "Effect of Party on Topic Prevelance for Facebook Posts of NC State Senators",
                                  xlim = c(-.08,0.08))
```

```{r}
# time
par(mfrow = c(2,2),mar = c(4,4,2,2))
for (i in 1:k){
  plot.estimateEffect(prep, "Time", method = "continuous", topics = rank[i], model = z,  
                      main = paste0(topic$topicnames[i],": Topic ",i),
                      printlegend = FALSE, ylab = "Exp. Topic Prob", 
                      xlab = "Time (Month, 1 = Jan 2015 to 24 = Dec 2016)", ylim = c(-0.01,0.2)
                      )
}
```

#### Comparing HB2-related Topics

```{r}
# Bathroom Safety, HB2
par(mfrow = c(1,1),mar = c(4,4,2,2))
plot.STM(stmFit, type = "perspectives", topics = c(26,16), n = 30, plabels = c("Bathroom Safety","HB2"))
```

#### Comparing Governor-related Topics

```{r}
# Pat McCrory and Roy Cooper
plot.STM(stmFit, type = "perspectives", topics = c(6,33), n=30, plabels = c("Pat McCrory","Roy Cooper"))
```

#### Comparing Presidential Election

```{r}
# Pres Election and Hillary Clinton
plot.STM(stmFit, type = "perspectives", topics = c(11,29), n=30, plabels = c("Presidential Election","Hillary Clinton"))
```

### Topic Correlations

Let's create an interactive network for the topics (nodes) in which an edge represents a significant correlation between the topic. The size of the node is the prevalence (expected topic proportion) for the topic.

```{r}
library(igraph); library(visNetwork)
par(mfrow = c(1,1))

mod.out.corr <- topicCorr(stmFit, cutoff = .025)

#library(corrplot)
#corrplot(mod.out.corr$cor, order="hclust", hclust.method="ward.D2", method = "circle", type = "lower", diag = F)

#mod.out.corr <- topicCorr(stmFit, method = "huge")

links2 <- as.matrix(mod.out.corr$posadj)
net2 <- graph_from_adjacency_matrix(links2, mode = "undirected")
table(V(net2)$type)
net2 <- simplify(net2, remove.multiple = F, remove.loops = T) 

links <- as_data_frame(net2, what="edges")
nodes <- as_data_frame(net2, what="vertices")

# Community Detection
clp <- cluster_label_prop(net2)
nodes$community <- clp$membership

means <- as.data.frame(unlist(STMresults$means))
colnames(means) <- "means"
color <- colorRamp(c("white","blue"))(abs(means$means)/0.05)
means$colorDem <- rgb(color[,1],color[,2],color[,3],  maxColorValue=255)

color <- colorRamp(c("white","red"))(abs(means$means)/0.05)
means$colorRep <- rgb(color[,1],color[,2],color[,3],  maxColorValue=255)

means$color <- ifelse(means$means>0,means$colorDem,means$colorRep)

#visNetwork edits
nodes$shape <- "dot"  
nodes$shadow <- TRUE # Nodes will drop shadow
nodes$title <- apply(topicNames$prob, 1, function(x) paste0(x, collapse = " + "))[rank] # Text on click
nodes$label <- topic$topicnames # Node label
nodes$size <- (topic$TopicProportions / max(topic$TopicProportions)) * 40 # Node size
nodes$borderWidth <- 2 # Node border width

nodes$color.background <- means$color
nodes$color.border <- "black"
nodes$color.highlight.background <- "orange"
nodes$color.highlight.border <- "darkred"
nodes$id <- topic$TopicNumber

visNetwork(nodes, links, width="100%",  height="600px", main="NC State Senator Topic (Correlation) Network") %>% visOptions(highlightNearest = list(enabled = TRUE, algorithm = "hierarchical")) %>% 
visInteraction(navigationButtons = TRUE) 
```


